---
title: "PA1_template"
author: "Guillem Mitjà"
date: "9/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read the activity data
By this code, the file activity.csv is read and the dates converted to the proper date type
```{r}
Activity <- read.csv("activity.csv")
Activity$date <- as.Date(as.character(Activity$date),"%Y-%m-%d")
```


# Calculation of the total number of steps per day
Now, the total number of steps per day is calculated and plot in the following histogram 
```{r echo=FALSE,results='hide',fig.keep='all'}
StepsDay <- with(na.omit(Activity), tapply(steps, date, sum))
par(mfrow = c(1,1), mar = c(1,1,1,1))
hist(StepsDay, breaks = 30, main = "Histogram total daily steps")
dev.copy(png, file = "figures/plot1.png")
dev.off()
```

# Mean and median of steps per day
The calculated mean of the steps per day is `r format(mean(StepsDay), digits = 7)` steps. The median is `r median(StepsDay)` steps.


# Calculate Daily pattern

The following plot shows an averaged daily pattern for each interval of five minuts accross a day.

```{r echo=FALSE,results='hide',fig.keep='all'}
StepPattern <- with(na.omit(Activity), tapply(steps, interval, mean))
plot(unique(Activity$interval),StepPattern, type = "l", main = "Daily average pattern")
dev.copy(png, file = "figures/plot2.png")
dev.off()
max(StepPattern)
```


# Missing values adjustment and analysis

The total number of missing values is `r sum(is.na(Activity$steps))`. We can see how they compare to the values that are no na.

```{r echo = FALSE, results = 'hide'}
TabNA <- table(is.na(Activity$steps))
names(TabNA) <- c("Good values", "NA values")
```

```{r echo=FALSE}
knitr::kable(
  TabNA
)
```

# Filling the gaps

Using this code, the values that are NA are filled with the mean value for that specific interval for all the recorded measurements.

```{r}
ActivityNoNa <- Activity
for (i in 1:length(Activity$steps)) {
  if (is.na(Activity$steps[i])){
    if (i%%288 == 0){
      ActivityNoNa$steps[i] <- StepPattern[[288]]
    }else{
      ActivityNoNa$steps[i] <- StepPattern[[i %% 288]]}
    } else{
      ActivityNoNa$steps[i] <- Activity$steps[i]
    }
}
```


## Calculate mean and median

Once we have filled the NA gaps, the histogram of the total steps per day is generated again to compare the values after the NA removal.

```{r echo=FALSE,results='hide',fig.keep='all'}
StepsDayNoNa <- with(ActivityNoNa, tapply(steps, date, sum))
hist(StepsDayNoNa, breaks = 30, main = "Total Steps without n/a")
dev.copy(png, file = "figures/plot3.png")
dev.off()
```

The calculated mean of the steps per day after NA removal is `r format(mean(StepsDayNoNa), digits = 7)` steps. The median is `r format(median(StepsDayNoNa), digits = 7)` steps.
The following table shows the comparison

```{r echo=FALSE,results='hide'}
MatSteps <- matrix(c(mean(StepsDay), median(StepsDay),mean(StepsDayNoNa),median(StepsDayNoNa)) ,nrow = 2, ncol = 2)
rownames(MatSteps) <- c("mean", "median")
colnames(MatSteps) <- c("Original", "NA removed")
```
```{r echo=FALSE}
knitr::kable(
  MatSteps
)
```
The results are quite similar. Only the mean has been slightly affected after the NA correction. 

# Weekdays and weekend comparison

The first step is to add the type of day at the Activity dataset. The following code does it.

```{r results='hide'}
for (i in 1:length(ActivityNoNa$steps)) {
  if(weekdays(ActivityNoNa$date[i]) == ("dissabte") | weekdays(ActivityNoNa$date[i]) == ("diumenge")){
    ActivityNoNa$TypeDay[i] <- "weekend"
  }
  else {
    ActivityNoNa$TypeDay[i] <- "weekday"
  }
}
ActivityNoNa$TypeDay <- as.factor(ActivityNoNa$TypeDay)
```

# Plot Weekend and weekday 

The following plot compares the average of steps in a Weekday and in a weekend.

```{r echo=FALSE,results='hide',fig.keep='all'}
StepPatternWeekend <- with(ActivityNoNa[ActivityNoNa$TypeDay == "weekend",], tapply(steps, interval, mean))
StepPatternWeekday <- with(ActivityNoNa[ActivityNoNa$TypeDay == "weekday",], tapply(steps, interval, mean))
par(mfrow = c(2,1), mar=c(4,4,0.2,0.2))
plot(unique(Activity$interval),StepPatternWeekend,  type = "l", xlab = "Interval", ylab = "Steps Weekend")
plot(unique(Activity$interval),StepPatternWeekday,  type = "l", xlab = "Interval", ylab = "Steps Weekday")
dev.copy(png, file = "figures/plot4.png")
dev.off()
```




